An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-28 draft.tex
First Experiment Intro

Figure \ref{fig:experiment_flow} shows the workflow users can follow to run
experiments in Evo\-Swarm. The first step is to develop the search strategy
algorithms that are going to be used by the multi-population algorithm.
Algorithm developers can upload the container definition or images directly to
DockerHub; another option is to upload the container image definition (Dockerfile) to
GitHub and define a trigger to re-upload the image to DockerHub; this
can be defined in DockerHub (or any other registry) too.  Docker Compose
configuration files can reference these containers. When a user needs to deploy
the application, he or she needs to edit the Fi file
% Fi file? - JJ+
to specify the
type and number of containers needed for the application.  The {\tt docker-compose}
application is responsible for starting all the containers, downloading new
versions if it is necessary. After starting the application, users can push
Experiment Definition files and start monitoring the execution in the terminal.
After the experiment is finished, the user must execute another script
that takes as argument the
experiment id to generate a collection of files containing all the
data generated by the experiment in a JSON format.
Researchers can use these files to analyze and plot experiment data.  Python
scripts are included in the repository to plot the running times for the
experiments, and all other plots used in this paper;  there is also a script to
generate the files needed by the COCO framework \cite{hansen2016coco}, which can generate standard
comparisons against other methods.

@@

\begin{itemize}
 \item Whether we can improve the execution time of the algorithm by adding populations
 and serverless functions to the system. In particular, what is the effect of
 these changes on the scalability of the system?

@@

% Experiment conf

The EvoSwarm application uses configuration files to run noiseless BOBB testbed
experiments.  As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The \em{GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The \em{Instances} parameter indicates how many instances of each
function will be tested. \em{Instances}  has a defualt value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the \em{Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of  Function Evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the \em{Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

@@

>>>>

Figure \ref{fig:experiment_flow} shows the workflow users can follow to run
experiments in Evo\-Swarm. The first step is to develop the search strategy
algorithms that are going to be used by the multi-population algorithm.
Algorithm developers can upload the container definition or images directly to
DockerHub; another option is to upload the container image definition (Dockerfile) to
GitHub and define a trigger to re-upload the image to DockerHub; this
can be defined in DockerHub (or any other registry) too.  Docker Compose
configuration files can reference these containers. When a user needs to deploy
the application, he or she needs to edit the {\tt docker-compose.yml} file
% Fi file? - JJ+
% fixed - Mario
to specify the type and number of containers needed for the application.  The {\tt docker-compose}
application is responsible for starting all the containers, downloading new
versions if it is necessary. After starting the application, users can push
Experiment Definition files and start monitoring the execution in the terminal.
After the experiment is finished, the user must execute another script
that takes as argument the
experiment id to generate a collection of files containing all the
data generated by the experiment in a JSON format.
Researchers can use these files to analyze and plot experiment data.  Python
scripts are included in the repository to plot the running times for the
experiments, and all other plots used in this paper;  there is also a script to
generate the files needed by the COCO framework \cite{hansen2016coco}, which can generate standard
comparisons against other methods.

@@

\begin{itemize}
 \item Whether we can improve the execution time of the algorithm by adding populations
 and serverless functions to the system, in particular, what is the effect of
 these changes on the scalability of the system?

@@

% Experiment conf
The EvoSwarm application uses configuration files to run noiseless BOBB testbed
experiments.  As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The \em{GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The \em{Instances} parameter indicates how many instances of each
function will be tested. \em{Instances}  has a defualt value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the \em{Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of  Function Evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the \em{Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

@@

As we mentioned before, reactive systems can scale by adding additional copies
of serverless functions. In our case, we can start additional worker containers
to have the same effect. Regarding GAs using a similar architecture, the work of
Salza et al., suggest that worker nodes, need to have at least a certain level
of complexity, in terms of execution time,  to effectively scale on multiple
nodes, tending to linear scalability. In a population-based algorithm like
EvoSwarm, several parameters can increase or decrease the workload of workers.:

\begin{itemize}
    \item The number of populations: if there are more workers than populations, workers
    must wait for work to arrive. If there are too many populations, they could be
    standing in the queue for more time. 
    \item The size of populations and the number of generations: 
    These parameters naturally increase the execution time.
    \item The complexity of the algorithm: For instance, in our case, 
    the PSO implementation has a lower execution time than the GA.
\end{itemize}

To evaluate the effect of these parameters on the scalability of the system, we
propose the following experiment:
